
---
title: "Group 04 Notebook"
authors: Hannah Ziegler, Tinashe Chingarande, Andrés Roa
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [L.A. is slammed with record costs for legal payouts](https://www.latimes.com/local/lanow/la-me-ln-city-payouts-20180627-story.html#:~:text=Last%20budget%20year%2C%20the%20city,libraries%20or%20fixing%20its%20streets.)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

all_payouts_decade <- read_csv('data/finalized_clean_10_year_payouts.csv', skip=2, col_names =c("fiscal_year", "case_name", "claim_case_number", "amount", "department", "department_clean", "case_type"))

total_yearly_payouts <- read_csv('data/cleaned_total_yearly_payouts.csv')

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Beyond the mammoth cases, a pileup of other large payouts has padded the tab, a Times analysis of city payouts found. While the total number of payouts has dipped over time, the average amount has increased.”

* **Analysis summary**: We found the number of payouts that were issues in fiscal year 2017. Then, we created a separate dataframe with all of the payouts from the decade before. We counted the number of payouts from the decade before. Then, we made a dataframe with the number of payouts from each year. After that, we calculated the percent change between 2007 and 2017. We also found the average amount of money paid in each payout for 2006 and 2016, then found the percent change. 

```{r}
# Put code to reverse engineer sentence here

payouts_2017 <- all_payouts_decade %>%
  filter(fiscal_year == 2017)

# Write the code to create a dataframe of payouts from 2006-2016 

payouts_2006_2016 <- all_payouts_decade %>%
  filter(fiscal_year < 2017) 

# Sum the total number of payouts from 2006-2016 (satisfies "five times as many as a decade before")

payouts_2006_2016 %>%
  summarise(
    count=n())

# Determine the percent change in payouts through the whole 10-year period (satisfies "while the total number of payouts has dipped over time")

yearly_payout_count <- all_payouts_decade %>%
  group_by(fiscal_year) %>%
  summarise(count=n()) %>%
  arrange(fiscal_year)
  
yearly_payout_count %>%
  summarise(
  total_payouts_10_years = sum(count)
)

yearly_payout_count %>%
  summarise(
  percent_change = ((1369-1468)/18703)*100
)

# Find the average amount of payouts from the decade 2006-2016 (satisfies "the average amount has increased")

payouts_2006_2016 %>%
  summarise(
  total_mean_pay_amount = mean(amount)
  )

payouts_2006_2016 %>%
  filter(fiscal_year == 2006) %>%
  summarise(
  mean_pay_amount_2006 = mean(amount)
  )

payouts_2006_2016 %>%
  filter(fiscal_year == 2016) %>%
  summarise(
  mean_pay_amount_2016 = mean(amount)
  )
  
yearly_payout_count %>%
  summarise(
  percent_change = ((94230.15-25345.32)/37567.26)*100)

```


# Display results of code below this codeblock

### Sentence 2

* **Sentence text**: "Last budget year, the city paid out 30 settlements of $1 million or more — five times as many as a decade before."
* **Analysis summary**: We used our dataframe of all 2017 payouts to filter only for payout amounts that are greater than $1 million. Then, we found the sum of the total number of payouts greater than $1 million from the year 2007 to determine if the 30 payouts in 2017 were 5x greater than a decade before. 

```{r}
# Put code to reverse engineer sentence here

# Filter to find the total number of payouts over $1 million in 2017. (satisfies "Last budget year, the city paid out 30 settlements of $1 million or more")

payouts_2017 %>%
  filter(amount >= 1000000) 

# Sum the total number of payouts greater the $1 million from 2006-2016 (satisfies "five times as many as a decade before")

payouts_2006_2016 %>%
  filter(fiscal_year == 2007 & amount >= 1000000) %>%
  select(fiscal_year, amount, case_name)

```

### Sentence 3

* **Sentence text**: That includes more than $24 million paid to two men who were freed after decades in prison, along with lawsuits over arrests, injuries, car crashes and alleged retaliation and discrimination against its own employees.”

* **Analysis summary**: We researched another Los Angeles Times article (https://www.latimes.com/local/lanow/la-me-ln-wrongful-convictions-20160119-story.html) that identified the two men that receive the payouts that totaled to 24 million. We filtered the data to narrow the police_department_payouts_2016 dataframe to only include payments about $7 million. Then, we filtered out the single other case that fit these parameters. Then, we summarized the amount of these two payouts to confirm that it was indeed $24 million. To find the second half of this sentence, we made a dataframe of police payouts by the type of case and summarized the number of cases present for each type, thus confirming that there have been lawsuits over arrests, injuries, car crashes and alleged discrimination against the department's own employees which have padded the tab for LAPD's legal payouts in the last decade.

```{r}
# Put code to reverse engineer sentence here

police_department_payouts_2016 <- all_payouts_decade %>%
  filter(department_clean == "Police Department" & fiscal_year == 2016 & amount > 7000000)
  
  
police_department_payouts_2016 %>%
  filter(claim_case_number != "BC453870") %>%
  summarize(total = sum(amount))

police_payouts <- all_payouts_decade %>%
  group_by(case_type) %>%
  filter(department_clean == "Police Department") %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: “Thousands of legal battles involving the Los Angeles Police Department, including a rash of recent settlements over police shootings and wrongful imprisonment, have added up to more than 40% of the total — a higher sum than any other city department, the analysis showed." 

* **Analysis summary**: We confirmed that the department's total payouts were greater than any other department in the city by summarizing the amount of payouts that each other department had and calculating the percent of the total city payouts that these expenditures accounted for. We were able to confirm that police department payouts accounted for more than 40% of the total payouts across the decade. 

```{r}
# "Thousands of legal battles" (using payouts as a proxy)

all_payouts_decade %>%
  summarise(
    count=n())

# Total amount of payouts over 10 years = 880687820

# Total Police Department Sum over 10 years

all_payouts_decade %>%
  filter(department_clean == "Police Department") %>%
  group_by(department_clean) %>%
  summarize(total = sum(amount))

# Percentage of payouts that have been attributed to the police department across the last decade

(369018871)/880687820*100

all_payouts_decade %>%
  filter(department_clean != "Police Department") %>%
  group_by(department_clean) %>%
  summarize(total = sum(amount)) %>%
  mutate(percent = total/880687820*100) %>%
  arrange(desc(total))

```


### Sentence 5

* **Sentence text**: "In recent years, the city agreed to pay $15 million to end a legal battle over a man who was gravely injured while crossing Franklin Avenue in Hollywood, another $15 million to the family of two girls hit in a North Hollywood crosswalk and $9.5 million to the family of a teenage girl fatally struck by a taxi near Dockweiler State Beach."

* **Analysis summary**: 

```{r}

all_payouts_decade %>%
  filter(amount == 15000000)

all_payouts_decade %>%
  filter(amount == 9500000)

all_payouts_decade %>%
  filter(department_clean == "Department of Transportation") %>%
  filter(str_detect(case_name, 'Leopold'))

all_payouts_decade %>%
  filter(department_clean == "Department of Transportation") %>%
  filter(str_detect(case_name, 'Larsen'))

# payouts_2018 <- all_payouts_decade %>%
#  filter(fiscal_year == 2018)

# payouts_2017 %>%
#  summarize(total = sum(amount))

```